Simplifying Lotka--Volterra interaction matrices by species grouping: optimal groupings are unpredictable

### 2022-05-22 article.rnw
Use \cref for an equation reference.
Standardise the format of equation labels to that used for figures.
Add a sentence regarding the success of data-informed grouping in networks.

We propose that species might be best partitioned into groups without regard to any \emph{a priori} definition but instead based purely upon which grouping gives the most parsimonious model.
% DBS: We can say something here about how this data-informed approach has worked very well for other clustering problems, such as modularity in networks. I can point to some very good references along those lines.
In this study, we investigate the usefulness and correctness of such an approach using experimental data from competition between different \emph{Drosophila} species provided by \textcite{terry2021natural}.
We fit a Lotka--Volterra model to these data to find the groupings that give the most parsimonious model, and compare these groupings to similarity based upon a species trait (body size).
While doing so, we group species independently with regard to their effects upon other species and their responses to other species, to find whether the data-informed groupings differ between these two cases.

@@

The data provided the change in population size from one generation to the next, so we used a discrete-time version of the Lotka--Volterra competition model given by:
\begin{equation}
	N_{i,t+1} - N_{i,t} = N_{i,t} \, r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right) \,,
	\label{LVmodel}
\end{equation}
where $N_{i,t}$ is the population of species $i$ at time $t$ (the size of the parent generation), $N_{i,t+1}$ is the population at time $t+1$ (the size of the offspring generation), $r_i$ is the intrinsic growth rate of species $i$, and $\alpha_{ij}$ is the per capita effect of species $j$ on species $i$.
The competition coefficients were not constrained to be positive, allowing for the possibility of mutualistic effects between some species.
With $n$ species, this model has $n+n^2$ parameters: $n$ growth rates and $n^2$ competition coefficients.

@@

The Lotka--Volterra competition model in Eq.~(\ref{LVmodel}) is not linear in the parameters.
With the rows of the interaction matrix ungrouped, the regression could be reparameterised to be linear in the parameters by treating $r_i \alpha_{ij}$ as a single regression parameter for each $i,j$ pair.
With the rows grouped, however, this becomes impossible since each species still has an independent growth rate.
As such, we instead fit the models with the Gauss--Newton method for non-linear least squares and assumed a normal error distribution of the response variable.
To prevent heteroskedasticity and improve normality, we transformed total population growth to per capita population growth, which corresponds to the following model:
\begin{equation}
	\frac{N_{i,t+1} - N_{i,t}}{N_{i,t}} = r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right) \,.
\end{equation}

>>>>

We propose that species might be best partitioned into groups without regard to any \emph{a priori} definition but instead based purely upon which grouping gives the most parsimonious model.
Data-informed grouping approaches have previously proved successful in other contexts, such as finding modules in networks \parencite{guimera2005functional, newman2006modularity, girvan2002community}.
In this study, we investigate the usefulness and correctness of such an \emph{a posteriori} approach using experimental data from competition between different \emph{Drosophila} species provided by \textcite{terry2021natural}.
We fit a Lotka--Volterra model to these data to find the groupings that give the most parsimonious model, and compare these groupings to similarity based upon a species trait (body size).
While doing so, we group species independently with regard to their effects upon other species and their responses to other species, to find whether the data-informed groupings differ between these two cases.

@@

The data provided the change in population size from one generation to the next, so we used a discrete-time version of the Lotka--Volterra competition model given by:
\begin{equation}
	N_{i,t+1} - N_{i,t} = N_{i,t} \, r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right) \,,
	\label{eq:lv-model}
\end{equation}
where $N_{i,t}$ is the population of species $i$ at time $t$ (the size of the parent generation), $N_{i,t+1}$ is the population at time $t+1$ (the size of the offspring generation), $r_i$ is the intrinsic growth rate of species $i$, and $\alpha_{ij}$ is the per capita effect of species $j$ on species $i$.
The competition coefficients were not constrained to be positive, allowing for the possibility of mutualistic effects between some species.
With $n$ species, this model has $n+n^2$ parameters: $n$ growth rates and $n^2$ competition coefficients.

@@

The Lotka--Volterra competition model in \cref{eq:lv-model} is not linear in the parameters.
With the rows of the interaction matrix ungrouped, the regression could be reparameterised to be linear in the parameters by treating $r_i \alpha_{ij}$ as a single regression parameter for each $i,j$ pair.
With the rows grouped, however, this becomes impossible since each species still has an independent growth rate.
As such, we instead fit the models with the Gauss--Newton method for non-linear least squares and assumed a normal error distribution of the response variable.
To prevent heteroskedasticity and improve normality, we transformed total population growth to per capita population growth, which corresponds to the following model:
\begin{equation}
	\frac{N_{i,t+1} - N_{i,t}}{N_{i,t}} = r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right) \,.
\end{equation}
